---
title: "Live NBA Predictions"
author: "Ramzy Al-Amine"
date: "11/19/2020"
output: 
  html_document:
    css: style.css
---

```{r setup, echo = FALSE, include=FALSE}

library(tidyverse)
library(nbastatR)
library(tidymodels)

# folder setup
setwd("~/R/nba_forecasting")
data_folder <- "data"
src_folder <- "src"
output_folder <- "output"

# relevant files
model_file <- file.path(src_folder, "PredictionModel.rds")

# load functions 
source(file.path(src_folder, "functions.R"))

```

### Summary

I use random forests to forecast win probabilities for NBA teams, based on performance over the 5 previous games. 

### Model Description 

Performance is measured by advanced stats such as net rating, percent assists, percent turnovers, and a bunch of others (I will be adding info soon). I consider lags of 1, 2, and 3 games as well as a moving average of the previous 3 and 5 games to account for a team's momentum relative to its opponent going into a game. Considering all the lag transformations for both home and away team, the number of variables surpass 300. For now, the model is trained on 2018-19 data only.

### Expected Accuracy  

Right now, the model predicts the correct outcome exactly 2 thirds of the time. I expect this improve once I add training data and expand the features.

### Code

The code is written in R and is available on my Github. 

### Today's date(*)

For now, predicitons are made for games played on this day *last* year
```{r dataset, echo = FALSE, include = FALSE}

# set date for which to predict games
this_day <- Sys.Date() - 365
this_season <- 2020 # indicates 2019-20 season

game_ids <- game_logs(seasons = this_season, result_types = "team", season_types = "Regular Season") %>%
  filter(dateGame <= this_day, dateGame >= this_day - 45) %>% 
  select(idGame) %>% 
  unique()

# transform a series of Game IDs into modeling dataste
new_games <- prep_features(game_ids, this_season) %>%
  mutate_if(is.logical, as.factor) %>%
  mutate_if(is.character, as.factor)

this_day

```


```{r predictions, echo=FALSE}

# load predictive model
pred_model <- readRDS(model_file)

# predict on new data set
predictions <- pred_model %>%
  predict(new_games, type = c("class"))

probabilities <- readRDS(model_file) %>%
  predict(new_games, "prob")

pred_df <- new_games %>%
  select(idGame, dateGame, isWin, slugTeam, slugOpponent) %>%
  bind_cols(predictions) %>%
  bind_cols(probabilities) %>%
  mutate(correct = case_when(isWin == .pred_class ~ "Correct", TRUE ~ "Incorrect"))

pred_df

```


```{r}

# analyze overall performance
pred_df %>%
  count(correct) %>%
  mutate(pct = n / sum(n))

# analyze performance for close calls
close_calls <- (.25:.75)
pred_df %>%
  filter(between(.pred_TRUE, 0.25, 0.75)) %>%
  count(correct) %>%
  mutate(pct = n / sum(n))

plot_accuracy(pred_df)

```

